A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species

### 2016-10-11 doc/paper1-submission/paper4.tex
Use non-numbered pipeline figure.
Refer to out tree as OTLv4' to save space.
Update and add second table for representing taxonomy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}
\newcommand{\otl}{OTLv4}
\newcommand{\otlprop}{OTLv4/propinquity}

@@

Propinquity was designed to function as a part of the Open Tree of
Life software architecture, so the first few steps of the pipeline
involve transforming artifacts from that project into a set of rooted
trees and a phylogenetic taxonomy. The phylesystem API \citep{McTavishEtAt2015}
of Open Tree allows users to curate published estimates of trees and
create ranked collections of these trees. Early steps in the propinquity
pipeline manipulate the phylogenetic input trees to improve their
usability and reliability. The first steps of the pipeline (see Figure
\ref{fig:pipeline}) collect a list of trees to include (in the \texttt{phylo\_input}
subdirectory) and store copies of these files (in the \texttt{phylo\_snapshot}
subdirectory) to make it easier to replicate the operation (because
the collection of trees and the tree files change due to curation).
\begin{figure}
\begin{centering}
\includegraphics[height=0.75\paperheight]{paper-pipeline-tools.pdf}
\par\end{centering}
\caption{Organization of the \emph{propinquity} pipeline. Each colored pentagon
labels a program (blue for otcetera-based tools and red for python
scripts in the propinquity or peyotl repository) that performs the
important operations in each step; the number before the tool name
refers to the section in this paper that describes the operation.
The output of each step corresponds to a subdirectory of the propinquity
system which will hold the output artifacts for the step. Ovals are
resources that are required (OTT and Open Tree's phylesystem repository).
White pentagons are user-controlled inputs.}
\label{fig:pipeline}
\end{figure}

@@

We seek to assess the performance of our new supertree method by
comparing it to the supertree method of \cite{HinchliffEtAl2015}.  The
method of \cite{HinchliffEtAl2015} was used to construct the Open Tree
of Life v4 (\otl{}).  Therefore, in order to facilitate comparison, we 
applied our method to same input trees and taxonomy used by
\otl{}.  We used the same cleaning flags to trim the taxonomy and
input trees.  We refer to the resulting supertree as \otlprop{} since it was
constructed by applying the propinquity pipeline to the same inputs as
\otl{}.

@@

The flag-cleaned version of OTT contained 2,424,255 leaves.  The
\otl{} supertree was constructed from 482 phylogenetic inputs,
containing a total of 45385 leaves, of which 41029 were unique.
After flag-cleaning and exemplification by propinquity, these trees contained 40323
unique tips.

@@

% displayed in v4: 35,518
% displayed in v4propr: 39,639

@@

\subsection{splits}

\begin{tabular}{lcccccc}
 & supported\_by & partial\_path\_of & terminal & conflicts\_with & resolves & resolved\_by\tabularnewline
\hline 
OTLv4 & 34,595 & 923 & 45,385 & 2760 & 473 & 2718\tabularnewline
\hline 
\textendash{} only & 745 & 54 & 0 & 2055 & 0 & 2718\tabularnewline
\hline 
OTLv4/propinquity & 38,521 & 1,118 & 45,385 & 1,357 & 515 & 0\tabularnewline
\hline 
\textendash{} only & 4,671 & 249 & 0 & 652 & 42 & 0\tabularnewline
\hline 
\end{tabular}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}
\newcommand{\otl}{OTLv4}
\newcommand{\otlprop}{OTLv4$^\prime$}

@@

Propinquity was designed to function as a part of the Open Tree of
Life software architecture, so the first few steps of the pipeline
involve transforming artifacts from that project into a set of rooted
trees and a phylogenetic taxonomy. The phylesystem API \citep{McTavishEtAt2015}
of Open Tree allows users to curate published estimates of trees and
create ranked collections of these trees. Early steps in the propinquity
pipeline manipulate the phylogenetic input trees to improve their
usability and reliability. The first steps of the pipeline (see Figure
\ref{fig:pipeline}) collect a list of trees to include (in the \texttt{phylo\_input}
subdirectory) and store copies of these files (in the \texttt{phylo\_snapshot}
subdirectory) to make it easier to replicate the operation (because
the collection of trees and the tree files change due to curation).
\begin{figure}
\begin{centering}
\includegraphics[height=0.75\paperheight]{pipeline-tools.pdf}
\par\end{centering}
\caption{Organization of the \emph{propinquity} pipeline. Each colored pentagon
labels a program (blue for otcetera-based tools and red for python
scripts in the propinquity or peyotl repository) that performs the
important operations in each step; the number before the tool name
refers to the section in this paper that describes the operation.
The output of each step corresponds to a subdirectory of the propinquity
system which will hold the output artifacts for the step. Ovals are
resources that are required (OTT and Open Tree's phylesystem repository).
White pentagons are user-controlled inputs.}
\label{fig:pipeline}
\end{figure}

@@

We seek to assess the performance of our new supertree method by
comparing it to the supertree method of \cite{HinchliffEtAl2015}.  The
method of \cite{HinchliffEtAl2015} was used to construct the Open Tree
of Life v4 (\otl{}).  Therefore, in order to facilitate comparison, we 
applied our method to the same input trees and taxonomy used by
\otl{}.  
We refer to the resulting supertree as \otlprop{} since it was
constructed by applying the propinquity pipeline to the same inputs as
\otl{}.

@@

The flag-cleaned version of OTT contained 2,424,255 leaves.  The
\otl{} supertree was constructed from 482 phylogenetic inputs,
containing a total of 45385 leaves, of which 41029 were unique.
After flag-cleaning and exemplification by propinquity, these trees contained 40323
unique tips. We used the same cleaning flags to trim the taxonomy and
input trees when constructing \otlprop{}, so \otlprop{} contains the
same number of leaves.

@@

% displayed in v4: 35,518
% displayed in v4prop: 39,639

@@

\subsection{Representing input splits}

\begin{table}

\caption{Representation of input splits in the \otl{} tree and the \otlprop{} tree.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
 & supported\_by & partial\_path\_of & terminal & conflicts\_with & resolves & resolved\_by\tabularnewline
\hline 
\otl{} & 34,595 & 923 & 45,385 & 2760 & 473 & 2718\tabularnewline
\hline 
\otl{} only & 745 & 54 & 0 & 2055 & 0 & 2718\tabularnewline
\hline 
\otlprop{} & 38,521 & 1,118 & 45,385 & 1,357 & 515 & 0\tabularnewline
\hline 
\otlprop{} only & 4,671 & 249 & 0 & 652 & 42 & 0\tabularnewline
\hline 
\end{tabular}}

\end{table}

\begin{table}

\caption{Representation of taxonomy splits in the \otl{} tree and the \otlprop{} tree.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
 & supported\_by & partial\_path\_of & terminal & conflicts\_with & resolves & resolved\_by\tabularnewline
\hline 
\otl{} & 125,384 & 0 & 2,424,255 & 1998 & 3676 & 4\tabularnewline
\hline 
\otl{} only & 296 & 0 & 0 & 19 & 17 & 4\tabularnewline
\hline 
\otlprop{} & 125,107 & 0 & 2,424,255 & 2,279 & 3,883 & 0\tabularnewline
\hline 
\otlprop{} only & 19 & 0 & 0 & 300 & 224 & 0\tabularnewline
\hline 
\end{tabular}}

\end{table}

We performed an annotation of the \otl{} tree to compare the ability
of the \otl{} tree and the \otlprop{} to represent splits in the input trees.
